---
title: "important_plots_analysis_ST_JBO002"
author: "aaron kollo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load libraries:

```{r, message=FALSE}
library(STutility)
library(dplyr)
library(tidyverse)
library(data.table)
library(reshape2)
library(readxl)
library(ggplot2)
library(Seurat)
library(RColorBrewer)
```
# load ST/SC data
```{r}
infoTable <- read.csv("livercellatlas_data/visium/mouse/infoTable.csv", header = TRUE, sep =",")
sample <- "JBO002"
sample.path <- paste("livercellatlas_data/visium/mouse", infoTable[infoTable$short_file_name == sample,]$geo_accession, sep = "/")
se <-  InputFromTable(infotable = infoTable[infoTable$short_file_name == sample,],
                      minUMICountsPerGene = 100,
                      minSpotsPerGene = 5,
                      minUMICountsPerSpot = 500,
                      platform =  "Visium")

img.save.path <- r"(C:\Users\aaronko\OneDrive - Karolinska Institutet\Uni\Master\4. Semester\thesis_project\writing\images\sliding_window\ST\JBO002\)"
```

```{r}
se <- LoadImages(se, time.resolve = FALSE)
```

==\> note that this is only applicable to 1 samples. for more samples it has to be modified

View image:

```{r}
ImagePlot(se, method = "raster")
```

Normalize data: se \<- SCTransform(se,return.only.var.genes = T)

```{r, message=FALSE, include=FALSE}
se <- SCTransform(se,return.only.var.genes = T)
```

FeatureOverlay

```{r, fig.width=8, fig.height=4}

spot_alpha = 0.2
point.size = 1.2

FeatureOverlay(se, 
               features = c("nCount_RNA","nFeature_RNA"),
               pt.alpha = spot_alpha,
               pt.size = point.size,
               add.alpha=T,
               ncols = 2,
               cols = c("black", "darkblue", "cyan", "yellow", "red", "darkred"))
```

## load zone and annotated seurat_clusterss from "spatial_structures" file

```{r, load clusters and zones}
res <- 0.8
seurat_clusters <- read.csv(paste0(sample.path, "/seurat_clusters_res",res, ".csv"), sep ="\t", header = T)
se$seurat_clusters <- seurat_clusters
se$seurat_clusters <-as.factor(se$seurat_clusters)

seurat_zones_df <- read.csv(paste(sample.path,"seurat_clusters_annot_zones.csv",sep = "/"), sep ="\t", header = T) 
seurat_zones <- seurat_zones_df$seurat_clusters
names(seurat_zones) <- row.names(seurat_zones_df)
seurat_zones <- sort(seurat_zones)
se$zone <- as.factor(seurat_zones)
```

```{r}
library(viridis)
n <- 18
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
col_vector[6] <- "white"
col_vector[8] <- "lightpink"
# col_vector2 <- c("darkred","orange","lightblue","darkblue",col_vector)
n <- length(unique(se$seurat_clusters))
col_vector2 <- turbo(n = n, begin = 0.2, end = 1)
se <- SetIdent(se, value="seurat_clusters")
```

## Cluster Feature Overlay

```{r,fig.width=10, fig.height=4}
# col_vector2 <- c("orange","lightblue","red","darkblue")

p1 <- FeatureOverlay(object = se, features = "zone", pt.alpha = 0.6, pt.size = 1.25, cols = col_vector2, pt.border = FALSE)
p2 <- FeatureOverlay(object = se, features = "seurat_clusters", pt.alpha = 0.6, pt.size = 1.25, cols = col_vector2, pt.border = FALSE)
p3 <- FeatureOverlay(object = se, features = "Cyp2e1", pt.alpha = 0.6, pt.size = 1.25, cols = c("black", "darkblue", "cyan", "yellow", "red", "darkred"), pt.border = FALSE)

p1|p2
p2|p3

```

## figure 1b and 3b
```{r, fig.height=4, fig.width=5, eval = F, include = F}
p2 <- ST.FeaturePlot(object = se, features = "seurat_clusters", pt.size = 1.1, cols = col_vector2, pt.border = FALSE)+
    theme(
    # axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.key.width = unit(0.5, "cm"),
        legend.key.height = unit(0.5, "cm"))+
    guides(colour = guide_legend(override.aes = list(size=10)),
           fill = guide_legend(override.aes = list(size=3)))+
  labs(legend = "seurat clusters",
       fill = "seurat clusters")

p2

ggsave(filename = paste0(img.save.path, "seurat cluster res 0.8.svg"),
       plot = p2,
       device = "svg",
       dpi = 96,
       height = 4,
       width = 5)
```

```{r}
meta <- se[[]]%>%
  mutate(zone = ifelse(seurat_clusters %in% c(2,4,6), "pericentral",
                       ifelse(seurat_clusters %in% c(0,5), "mid",
                              "periportal")))%>%
  select(zone)
se <- AddMetaData(se, meta)
```

```{r, fig.height=4, fig.width=5, eval = F, include = F}
p2 <- ST.FeaturePlot(object = se, features = "zone", pt.size = 1.1, cols = col_vector2, pt.border = FALSE)+
    theme(
    # axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.key.width = unit(0.5, "cm"),
        legend.key.height = unit(0.5, "cm"))+
    guides(colour = guide_legend(override.aes = list(size=10)),
           fill = guide_legend(override.aes = list(size=3)))+
  labs(legend = "seurat clusters",
       fill = "seurat clusters")+
  scale_fill_viridis_d()

p2

```


```{r, fig.height=4, fig.width=5, eval = F, include = F}
ggsave(filename = paste0(img.save.path, "zone featureoverlay.svg"),
       plot = p2,
       device = "svg",
       dpi = 96,
       height = 4,
       width = 5)
```


# load data from enrichment analysis
```{r}
important_plots <- readRDS(file = paste0(sample.path, "/important_plots_",sample,".rds"))
  
  # readRDS(file = paste0(sample.path, "/important_plots_", sample,".rds")
```

```{r}
top_genes_df <- lapply(important_plots, function(x) x$top.genetic.genes.df) %>%
  bind_rows(.id = "disease")

```

```{r}
head(top_genes_df, 10)

```


```{r}
top_genes_df <- top_genes_df%>%
  mutate(significance.level = ifelse(p.value.adjusted < 0.001, "***",
                     ifelse(p.value.adjusted < 0.01, "**",
                     ifelse(p.value.adjusted < 0.05, "*",
                     ""))))
```

## plotting
### figure 3a
spatial p plot
```{r}
p.vals <- se[[]]
p.vals<- p.vals %>%
  left_join(top.genetic.genes) %>%
  select(!contains("rank"))
rownames(p.vals) <- rownames(se[[]])
p.se <- se 
p.se <- AddMetaData(p.se, p.vals)
p <- ST.FeaturePlot(object = p.se,
              features = "log10.p",
              pt.size = 1,
              cols = col_vector2
              )+
  labs(fill="-log10(p-value)", title = "") 
p
```

### geom_tile

```{r, fig.width=8, fig.height=10}
p <- top_genes_df %>%
  ggplot(aes(x = seurat_clusters, y = disease, fill = significant))+
  geom_tile(colour = "grey20")+
  scale_fill_manual(values = c("grey75","red"), breaks = c("FALSE", "TRUE"))+
  theme(
    # axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.key.width = unit(0.5, "cm"),
        legend.key.height = unit(0.5, "cm"))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 22))+
  coord_equal()+
  labs(x = "seurat cluster",
       fill = "significant Bonferroni\n adjusted p-value")

p
```

### figure 3c
enrichment analysis matrix
```{r, fig.width=8, fig.height=10}
tick.size <- 10
print(unique(top_genes_df$significance.level))
p <- top_genes_df %>%
  ggplot(aes(x = seurat_clusters, y = disease, fill = log10.p))+
  geom_tile(aes(color = significance.level))+
  geom_tile(colour = "grey20", aes(fill = log10.p))+
  # geom_point(aes(shape = significant))+
  geom_text(data = top_genes_df %>% filter(significant == TRUE),
              aes(label = significance.level),
              vjust = 0.75,
              size = 5,
              fontface = "bold"
              )+
  # scale_fill_manual(values = c("grey75","red"), breaks = c("FALSE", "TRUE"))+
  theme(
    # axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.key.width = unit(0.5, "cm"),
        legend.key.height = unit(0.5, "cm"),
        axis.text.x = element_text(face = "bold", size = 10),
        axis.text.y = element_text(face = "bold", size = 10))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 22))+
  coord_equal()+
  labs(x = "seurat cluster",
       fill = "-log10(p-value)")+
  scale_fill_viridis_c(begin = 0.2, option = "turbo")+
  scale_color_manual(values = c("***" = "white", "**" = "white"),
                     labels = c("***" = "***  p < 0.001", "**" = "**  p < 0.01"),
                     guide = guide_legend(override.aes = list(shape = c(21, 23),
                                                              fill = c("white", "white"),
                                                              size = 4)))

p

```

```{r}
ggsave(filename = paste0(img.save.path, "geom_tile log10 pvalue.svg"),
       plot = p,
       device = "svg",
       dpi = 96,
       height = 10,
       width = 8)
```


