---
title: "scRNA_SW_healthy_pvalue_CT_short"
author: "aaron kollo"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "reports")})
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    dev: 'CairoPNG'
date: "Sys.Date()"
---

Load libraries:

```{r, message=FALSE}
library(dplyr)
library(tidyverse)
library(data.table)
library(reshape2)
library(readxl)
library(ggplot2)
library(Seurat)
library(RColorBrewer)
library(SeuratDisk)
```

# setup
```{r setup, include=FALSE}
library(knitr)

# which plots to print:
violin <- F
lineplots <- T
p.plots <- T


# plot setups
# violin.w <- 12 
# violin.h <- 9
# 
# working with 4K monitor?
big_screen <- T

# global chunk settings
opts_chunk$set(echo = TRUE)
opts_chunk$set(dev = "CairoPNG")
opts_chunk$set(error = FALSE) # when FALSE, interrupt in case of errors                



opts_hooks$set(
# eval
  eval = function(options) {
    raw.code <- knit_code$get(opts_current$get()$label) 
    raw.code <- paste(raw.code, collapse = " ")
    if (grepl("p_value|p.val|top.genetic.genes", raw.code)) {
      options$eval = p.plots
      }
    if (grepl("<<line plot>>|line.plot.setup", raw.code)){
      options$eval = lineplots
      }
    if (grepl("<<Violin Plots>>|geom_violin", raw.code)){
      options$eval = violin
      }
  options
  })

opts_hooks$set(
# include
  eval = function(options) {
    raw.code <- knit_code$get(opts_current$get()$label) 
    raw.code <- paste(raw.code, collapse = " ")
    if (grepl("p_value|p.val|top.genetic.genes", raw.code)) {
      options$include = p.plots
      }
    if (grepl("<<line plot>>|line.plot.setup", raw.code)){
      options$include = lineplots
      }
    if (grepl("<<Violin Plots>>|geom_violin", raw.code)){
      options$include = violin
      }
  options
  })

opts_hooks$set(
# fig.width
  fig.width = function(options) {
    raw.code <- knit_code$get(opts_current$get()$label) 
    raw.code <- paste(raw.code, collapse = " ")
    if (grepl("p_value|p.val|top.genetic.genes", raw.code)) {
        options$fig.width = 12
        }
    if (grepl("line plot|line.plot.setup", raw.code)){
        options$fig.width = 30
        }
    if (grepl("<<Violin Plots>>|geom_violin", raw.code)){
        options$fig.width = 12 
        }

    options
  })
opts_hooks$set(
# fig.height
  fig.height = function(options) {
    raw.code <- knit_code$get(opts_current$get()$label) 
    raw.code <- paste(raw.code, collapse = " ")
    if (grepl("p_value|p.val|top.genetic.genes", raw.code)) {
      options$fig.height = 9
      }
    if (grepl("line plot|line.plot.setup", raw.code)){
      options$fig.height = 30
      }
    if (grepl("<<Violin Plots>>|geom_violin", raw.code)){
      options$fig.height = 9
      }
    options
  }
)

opts_hooks$set(fig.height = function(options) {
  if (big_screen == T) {
    options$fig.height = options$fig.height*1.5
  }
  options
})


```

# Load scRNA data:
```{r}
# se <- readRDS()
cond <- "healthy"
sample.path <-list.files(paste0("livercellatlas_data/RNAseq/", cond), full.names = T) %>%
                          .[grepl("rawData", .)] %>%
                          paste0(.,"/")

se <- LoadH5Seurat(paste0(sample.path, "sctransformed_pca_filtered_se.h5seurat"), assays = "SCT", reductions = FALSE, graphs = FALSE)
DefaultAssay(object = se)
se
length(Cells(se))
# se <- LoadH5Seurat(paste0(sample.path,"subset_filtered_se.h5seurat"))
```


# load other data

## load window meta data

from AddModuleScore (from before)

```{r}
list.files(sample.path)
meta.data <- fread(file = paste0(sample.path, "sliding_window_meta_data_5k.tsv"),
                      sep = "\t")
meta.data[c(1:5),c(1:20)]
colnames(meta.data)[1] <- "barcode"
meta.data <- meta.data %>%
  column_to_rownames(var = "barcode")

meta.data[c(1:10), c(1:10)]
# se <- AddMetaData(se, meta.data) # combine both dataframes on the rownames ==> remember, it needs cells as rownames
# is it needed for scRNA analysis? Maybe not, because you have the celltype column in the sliding window .tsv
```

```{r}
meta.data %>%
  select(celltype, contains("_rank_")) %>%
  colnames() %>%
  gsub("_rank_.*","",.)%>%
  unique()
```

```{r, make summary df of the windows}
# this step might be also done in uppmax and then further processed locally because the sliding window data is quite big, abou

summary.meta.clusters <- read.csv(file = paste0(sample.path, "sliding_window_meta_data_summarised_5k.tsv"),
sep = "\t")

library("psych")
#here you can decide whether you wanna summarise by cluster or by zone! 
#zone or seurat_clusters
summarise.on <- "celltype"
# # meta.data <- se[[]]
# 
# summary.meta.clusters <- meta.data %>%
#   dplyr::select(any_of(summarise.on), contains("_rank_"))
# 
# summary.meta.clusters <- describeBy(summary.meta.clusters, group = summary.meta.clusters[, summarise.on], quant = c(.05, .95)) 
# 
# summary.meta.clusters <- lapply(summary.meta.clusters, function(x) as.data.frame(x))
# 
# 
# summary.meta.clusters <- map(summary.meta.clusters, ~.x %>%
#                                rownames_to_column(., var = "window"))
# summary.meta.clusters <- bind_rows(summary.meta.clusters, .id = summarise.on)
# summary.meta.clusters <- summary.meta.clusters %>%
#   filter(!(window %like% "celltype"))
# # rownames(summary.meta.clusters) <- NULL
# head(summary.meta.clusters)
```

```{r}
genetic.summary.meta.clusters <- summary.meta.clusters %>%
  filter(window %like% "genetic")
drug.summary.meta.clusters <- summary.meta.clusters %>%
  filter(window %like% "drug")

```

## load NULL set data

```{r, load NULL gene set data}
list.files(sample.path)
NULL.set.data <- read.csv(paste0(sample.path, "se_null_set_summarised_5k.tsv"),
                          sep = "\t",
                          check.names = F) #summarized perm.matrix ==> median of all permutations per cluster

  
  
NULL.set.data <- describe(NULL.set.data, quant = c(.05, .95)) %>%
  as.data.frame %>%
  rownames_to_column(var = summarise.on) # summarise further using library(psych)
```




## introduce empty vectors to store data in

```{r for p value calculations, eval = F}
# not implemented yet, 11/04/23
top.genes.statistics <-  list()
```

# plotting

## load OT.files

is easier to select the top genes and plot the density/p-value plot

```{r, code_folding = 'hide'}
OT.files.split <- readRDS(file = paste("OpenTargets/scRNA", cond,"OT_files_split.rds", sep = "/"))

OT.files.split.genetic <- sapply(OT.files.split, function(x) x[grep("genetic", names(x), value = T)])
names(OT.files.split.genetic) <- names(OT.files.split)
OT.files.split.genetic <- OT.files.split.genetic[sapply(OT.files.split.genetic, function(x) length(x)>0)]


OT.files.split.drug <- sapply(OT.files.split, function(x) x[grep("drug", names(x), value = T)])
names(OT.files.split.drug) <- names(OT.files.split)
OT.files.split.drug <- OT.files.split.drug[sapply(OT.files.split.drug, function(x) length(x)>0)]
OT.files.split.drug <- sapply(OT.files.split.drug, function(x) x[[1]])
```

## NAFLD {.tabset}

subset features_to_plot to specific disease for now

```{r}
disease <- "NAFLD"
features_to_plot <- meta.data %>%
  select(celltype, contains(disease))

```

### Violin Plots

```{r Violin Plots, fig.height = 8, fig.width = 10}
#<<Violin Plots>>
p_list <- list()
for (col in colnames(features_to_plot)) {
  # Create a violin plot with boxplot and individual data points for the current column
  p <- features_to_plot%>%
    ggplot(aes(x = celltype, y = .data[[col]])) +
    geom_jitter(height = 0.1, width = 0.25, alpha = 0.2, size = 0.2) +
    geom_boxplot(width = 0.5, aes(fill = celltype), color = "black", alpha = 0.7) +
    geom_violin(trim = FALSE, aes(fill = celltype), alpha = 0.7) +
    labs(x = "celltype", y = col) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
  
  # plot p
  p_list[[col]] <- p
  plot(p)
}
```




#### line plot setup

```{r, class.source = 'fold-hide', echo = FALSE}
library(gridExtra)
library(cowplot)
#boxplot on the right side
p.boxplot <- NULL.set.data %>%
  ggplot(aes(x = celltype, ymin=min, lower= Q0.05, middle = median, upper = Q0.95, ymax = max)) +
  geom_boxplot(stat = "identity") +
  # ggtitle("")+
  theme_classic()+
  facet_wrap(vars(celltype), ncol = 1, scales = "free", strip.position = "left") #+
  # theme(
  # strip.background = element_blank(),
  # strip.text.x = element_blank()
  # )

#point/line plot on the left side
p.boxplot 
line.plot.setup <- function(x){
  lwd <- 0.2
  lty <- 2
  p1 <- ggplot(x, aes(window, median, group = 1)) +
  geom_point(shape = "o", size = 2) +
  geom_line(linewidth = 0.1) +
  geom_hline(data = NULL.set.data, aes(yintercept  = min), colour = "darkred", lwd = lwd, lty = lty)+
  geom_hline(data = NULL.set.data, aes(yintercept  = max), colour = "darkred", lwd = lwd, lty = lty)+
  geom_hline(data = NULL.set.data, aes(yintercept  = median), col = "blue", lwd = lwd, lty = lty)+
  geom_hline(data = NULL.set.data, aes(yintercept  = Q0.05),color = "red", lwd = lwd, lty = lty)+
  geom_hline(data = NULL.set.data, aes(yintercept  = Q0.95), colour = "red", lwd = lwd, lty = lty)+
  xlab("Rank") +
  ylab("Score") +
  scale_x_discrete(labels = c(1:nrow(x))) +
  theme_classic()+
  facet_wrap(~celltype, ncol = 1, scales = "free")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank())
  
  
  combined.plot <- plot_grid(p.boxplot, p1,
                             align = "h", axis = "b", ncol = 2, rel_widths = c(1, 20)) #https://stackoverflow.com/questions/41834093/decrease-margins-between-plots-when-using-cowplot, not used but can be adjusted
  combined.plot
}
```


### line plots

```{r line plot, fig.width = 30 , fig.height=30}
#<<line plot>>

genetic_features_to_plot <- summary.meta.clusters %>%
  filter(window %in% colnames(features_to_plot) & window %like% "genetic") 
drug_features_to_plot <- summary.meta.clusters %>%
  filter(window %in% colnames(features_to_plot) & window %like% "drug")

line.plot.setup(genetic_features_to_plot) 
# +
  # ggtitle(paste0("genetic associations ", disease))


# line.plot.setup(drug_features_to_plot) +
#   ggtitle("drug associations")

```

### p value permutation

#### top genetic genes

##### select top genes

```{r p_value}
nr.top.genetic.windows <- 12 #all
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
print(maximum)
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r p_value load_p_val_nullset}
p.val.nullset <- read.csv(paste0(sample.path, "/se_null_set_summarised_g", maximum, ".tsv"), sep = "\t", check.names = F) # i
colnames(p.val.nullset) <- gsub("X","", colnames(p.val.nullset))
stacked.p.val.nullset <- stack(p.val.nullset)
colnames(stacked.p.val.nullset) <- c("addmodulescore","celltype")
stacked.p.val.nullset$permutation <- rep(c(1:10000),length(unique(stacked.p.val.nullset$celltype)))
head(p.val.nullset)
```

##### addmodulescore for top genes

```{r p_value addmodulescore_top_genes, eval = p.plots}
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
top.genetic.genes <- AddModuleScore(se, features = list(top.genetic.genes), name = paste0(disease, "_top_genetic_genes")) %>%
  .@meta.data %>%
  select(contains(paste0(disease, "_top_genetic_genes")), celltype)

top.genetic.genes <- top.genetic.genes %>%
  group_by(celltype)%>%
  summarise(across(where(is.numeric), median))%>% 
  column_to_rownames("celltype") %>%
  t() %>%
  as.data.frame()%>%
  stack()
colnames(top.genetic.genes) <- c("median","celltype")

head(top.genetic.genes)
```

##### calculate p-value

```{r calculate p_value, eval = p.plots}
p.value <- imap(p.val.nullset, function(x, idx){
  print(idx) #this is the column name
  median.to.compare.against <-  top.genetic.genes[top.genetic.genes$celltype == idx, "median"]
  print(paste0("median: ", median.to.compare.against))
  length(x[x>=median.to.compare.against]) #for each column/celltype in the raw.NULL.set, subset the column to those rows that have a higher median than the median of the top (70) genes for each cluster
})



p.value <- sapply(p.value, function(x) (x+1)/(10000+1)) 
# calculate the fraction of the medians of the  10000 permutations that are higher than the median of the top.genes median 
# think about p-value definition:
# “the probability of obtaining test results at least as extreme as the results actually observed, under the assumption that the null hypothesis is correct.” 
p.value.adj <- p.adjust(p.value, method = c("bonferroni")) #, n=length(unique(top.genetic.genes$celltype))
p.value.adj <- data.frame("p.value.adjusted" = p.value.adj, "celltype" = as.factor(names(p.value.adj)))

top.genetic.genes <- top.genetic.genes %>%
left_join(p.value.adj)
top.genetic.genes <- top.genetic.genes %>%
  arrange(p.value.adjusted)%>%
  mutate(significant = ifelse(p.value.adjusted<0.05,T,F))

```

##### p-val plot setup

```{r p_value plot setup, eval = p.plots}
p.plot.setup <- function(top.genes.df){
  # adj_p_value <- ifelse(top.genes.df$p.value.adjusted<0.5,
  #                       top.genes.df$p.value.adjusted,
  #                       1-top.genes.df$p.value.adjusted)
  top.genes.df <- top.genes.df %>%
  arrange(p.value.adjusted)
  p <- ggplot(stacked.p.val.nullset, aes(x=addmodulescore)) +
  geom_density(fill = "#69b3a2",
               alpha = 0.8)+
  geom_vline(data = top.genes.df, aes(xintercept = median, color = significant),
             size = 1, show.legend = T) +
    scale_color_manual(values = c("yellow","red"))+
  geom_label(data = top.genes.df,
             aes(x=median+0.02, y = 8),
             label = paste0("adjusted p-value \n",
                      formatC(top.genes.df$p.value.adjusted,
                           format = "e", digits = 2)),
             size = 2) +
  xlab("Median score NULL gene set")+
  ylab("Density")
  
p +
  facet_wrap(~factor(celltype, levels = top.genes.df$celltype),
             ncol = 4)+
  theme_classic()+
  theme(legend.position="none")
}
```

### p-value density plot

```{r p_value plot, fig.height=9, fig.width=12, eval = p.plots}
p.plot.setup(top.genetic.genes)
```



## liver_disease_general {.tabset}

subset features_to_plot to specific disease for now

```{r}
disease <- "liver_disease_general"

features_to_plot <- meta.data %>%
  select(celltype, contains(disease))
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```



### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r, eval = p.plots}
nr.top.genetic.windows <- 14
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r, eval = p.plots}
<<p_value load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r, eval = p.plots}
<<p_value addmodulescore_top_genes>>
```

##### calculate p-value

```{r, eval = p.plots}
<<calculate p_value>>
```

### p-value density plot

```{r, eval = p.plots}
<<p_value plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster


## cirrhosis_of_liver {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "cirrhosis_of_liver"
features_to_plot <- meta.data %>%
  select(celltype, contains(disease))
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```


### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 10 #just all (107)
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<p_value load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<p_value addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p_value>>
```

### p-value density plot

```{r}
<<p_value plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

## cirrhosis_familial {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "cirrhosis_familial"
features_to_plot <- meta.data %>%
  select(celltype, contains(disease))

```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```



### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 9
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<p_value load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<p_value addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p_value>>
```

### p-value density plot

```{r}
<<p_value plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

## Diabetes type I {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "Diabetes_typeI"
features_to_plot <- meta.data %>%
  select(celltype, contains(disease))
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```



### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 18
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<p_value load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<p_value addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p_value>>
```

### p-value density plot

```{r}
<<p_value plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster


## Diabetes type II {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "diabetes_typeII"
features_to_plot <- meta.data %>%
  select(celltype, contains(disease))
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```



### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 11
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<p_value load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<p_value addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p_value>>
```

### p-value density plot

```{r}
<<p_value plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster


## triglyceride_measurement {.tabset}
subset features_to_plot to specific disease for now
OBS ==> there is also LDL_triglyceride_measurement, so have to exclude that 
```{r}
disease <- "triglyceride_measurement"

features_to_plot <- meta.data %>%
  select(celltype, contains(disease)& !contains("LDL"))
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```


### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 6
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<p_value load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<p_value addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p_value>>
```

### p-value density plot

```{r}
<<p_value plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster


## LDL_triglyceride_measurement {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "LDL_triglyceride_measurement"

features_to_plot <- meta.data %>%
  select(celltype, contains(disease))
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```



### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 24
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<p_value load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<p_value addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p_value>>
```

### p-value density plot

```{r}
<<p_value plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

## cholesterol_measurement {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "cholesterol_measurement"

features_to_plot <- meta.data %>%
  select(celltype, contains(disease) & !contains("LDL")& !contains("HDL")) 

```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```



### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 3
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<p_value load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<p_value addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p_value>>
```

### p-value density plot

```{r}
<<p_value plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

## LDL_cholesterol_measurement {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "LDL_cholesterol_measurement"

features_to_plot <- meta.data %>%
  select(celltype, contains(disease))
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```


### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 7
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<p_value load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<p_value addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p_value>>
```

### p-value density plot

```{r}
<<p_value plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster




## HDL_cholesterol_measurement {.tabset}
should be HDL_cholesterol
subset features_to_plot to specific disease for now
```{r, }
disease <- "HDL_cholesterol_measurement"

features_to_plot <- meta.data %>%
  select(celltype, contains(disease))
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```



### line plots

```{r, fig.width = 5, fig.height= 10}
<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 8
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<p_value load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<p_value addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p_value>>
```

### p-value density plot

```{r}
<<p_value plot>>
```


``` {r}
sessionInfo()
```