---
title: "LCA_preparation_sliding_window_JBO009"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    dev: 'CairoPNG'
date: "Sys.Date()"
---

Load libraries:
```{r, message=FALSE}
library(STutility)
library(dplyr)
library(tidyverse)
library(data.table)
library(reshape2)
library(readxl)
library(ggplot2)
library(Seurat)
library(RColorBrewer)
```


# setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)

big_screen <- T
knitr::opts_hooks$set(fig.width = function(options) {
  if (big_screen == T) {
    options$fig.width = options$fig.width*1.5
  }
  options
})
knitr::opts_hooks$set(fig.height = function(options) {
  if (big_screen == T) {
    options$fig.height = options$fig.height*1.5
  }
  options
})
knitr::opts_chunk$set(dev = "CairoPNG")
```

``` {r nullset setup}
nullset <- T #set if we want to calc null set or not

```

# Load ST data:
```{r}
infoTable <- read.csv("livercellatlas_data/visium/mouse/infoTable.csv", header = TRUE, sep =",")
sample <- "BH001"
sample.path <- paste("livercellatlas_data/visium/mouse", infoTable[infoTable$short_file_name == sample,]$geo_accession, sep = "/")
se <-  InputFromTable(infotable = infoTable[infoTable$short_file_name == sample,],
                      minUMICountsPerGene = 100,
                      minSpotsPerGene = 5,
                      minUMICountsPerSpot = 500,
                      platform =  "Visium")
# num_samples <- length(unique(se@meta.data[["condition"]]))
```



Load image:
```{r}
se <- LoadImages(se, time.resolve = FALSE)
```
==> note that this is only applicable to 1 samples. for more samples it has to be modified
```{r fig.width = 10, fig.height = 8}
p1 <- ggplot() +
  geom_histogram(data = se[[]], aes(nFeature_RNA), fill = "red", alpha = 0.7, bins = 50) +
  ggtitle("Unique genes per spot")

p2 <- ggplot() +
  geom_histogram(data = se[[]], aes(nCount_RNA), fill = "red", alpha = 0.7, bins = 50) +
  ggtitle("Total counts per spots")

gene_attr <- data.frame(nUMI = Matrix::rowSums(se@assays$RNA@counts), 
                        nSpots = Matrix::rowSums(se@assays$RNA@counts > 0))
p3 <- ggplot() +
  geom_histogram(data = gene_attr, aes(nUMI), fill = "red", alpha = 0.7, bins = 50) +
  scale_x_log10() +
  ggtitle("Total counts per gene (log10 scale)")

p4 <- ggplot() +
  geom_histogram(data = gene_attr, aes(nSpots), fill = "red", alpha = 0.7,  bins = 50) +
  ggtitle("Total spots per gene")

(p1 - p2)/(p3 - p4)

```

``` {r}
mt.genes <- grep(pattern = "^mt-", x = rownames(se), value = TRUE)
se$percent.mito <- (Matrix::colSums(se@assays$RNA@counts[mt.genes, ])/Matrix::colSums(se@assays$RNA@counts))*100
#theres a function for this: se$percent.mito <- PercentageFeatureSet(object = se, pattern = "^MT-")
#(“^MT-“) works for human gene names, "^mt-" for mouse!
# Collect all genes coding for ribosomal proteins
rp.genes <- grep(pattern = "^Rpl|^Rps", x = rownames(se), value = TRUE)
se$percent.ribo <- (Matrix::colSums(se@assays$RNA@counts[rp.genes, ])/Matrix::colSums(se@assays$RNA@counts))*100

se <- subset(se, percent.mito <= 15)

```




View image:
```{r}
ImagePlot(se, method = "raster")
```




Normalize data: se <- SCTransform(se,return.only.var.genes = T)
```{r, message=FALSE, include=FALSE}
se <- SCTransform(se,return.only.var.genes = T)
```

FeatureOverlay
```{r, fig.width=8, fig.height=4}

spot_alpha = 0.2
point.size = 1.2

FeatureOverlay(se, 
               features = c("nCount_RNA","nFeature_RNA"),
               pt.alpha = spot_alpha,
               pt.size = point.size,
               add.alpha=T,
               ncols = 2,
               cols = c("black", "darkblue", "cyan", "yellow", "red", "darkred"))
```

  
# Load OpenTarget files:
## function
```{r}
get_opentarget <- function (opentarget_dir = getwd(), mouse_genes = NULL ){
  OT.files <- list()
  disease.names <- character()
  i <- 1
  column.names <- c("Gene_symbol","Overall_assoc.score","Genetic_assoc.","Somatic_mutation","Drugs","Pathways_system_biology","Text_mining","RNA_expression","Animal_models","Target_name")
  
  if (!is.null(mouse_genes)){
    

      conversion.table <- read.table(mouse_genes,
                                     sep = "\t",
                                     header = TRUE) 
      #if mouse_genes dataframe is specified, load it

  }
  
  for (OT.file in list.files(opentarget_dir)){
    if (grepl("diseases_", OT.file)){
      print(OT.file)
  
      df <- read.table(paste(opentarget_dir, OT.file, sep= "/"),
                     sep= "\t",
                     header=TRUE,
                     quote = "")
    
      names(df) <- column.names  # rename columns
      
      df[df == "No data"] <- NA #Re-name N/A
    
      
    
      names(df) <- column.names  # rename columns
      
      df[df == "No data"] <- NA #Re-name N/A
      
      if (!is.null(mouse_genes)){
        df$Gene_symbol <- conversion.table[match(df$Gene_symbol,conversion.table$hgnc_symbol),]$mgi_symbol #convert human to mouse annotation
        OT.files[[i]] <- df
        } else {
          OT.files[[i]] <- df
        }
      i <- i+1
      disease.name <- sub(".*diseases_","", OT.file)
      disease.name <- gsub(".tsv", "", disease.name)
      disease.names <- append(disease.names, disease.name)
    }
  }

  names(OT.files) <-  disease.names
  OT.files <- map(OT.files, ~.x %>%
                    distinct())
  return(OT.files)
}
```

```{r}
conversion.t <- read.csv("OpenTargets/mouse_human_genes_feb_23.tsv", header = T, sep = "\t")
```

## loading files
```{r}
OpenTargets.path <- "OpenTargets"

OT.files <- get_opentarget( opentarget_dir = OpenTargets.path ,mouse_genes = "OpenTargets/mouse_human_genes.tsv")
OT.files.human <-get_opentarget(opentarget_dir = OpenTargets.path)

CVDs <- get_opentarget( opentarget_dir = "OpenTargets/CVDs" ,mouse_genes = "OpenTargets/mouse_human_genes.tsv")
CVDs.human <- get_opentarget( opentarget_dir = "OpenTargets/CVDs" )

lipid.diseases <- get_opentarget( opentarget_dir = "OpenTargets/lipids"  ,mouse_genes = "OpenTargets/mouse_human_genes.tsv")

lipid.diseases.human <- get_opentarget( opentarget_dir = "OpenTargets/lipids")

OT.files <- c(OT.files,  lipid.diseases)
OT.files.human <- c(OT.files.human, lipid.diseases.human)

```
## clean 
```{r, clean OT.files from duplicates}
OT.files <- map(OT.files, ~.x %>%
  dplyr::select(Gene_symbol, Overall_assoc.score, Genetic_assoc.,Drugs, Animal_models)%>%
  distinct() %>%
  filter(!is.na(Gene_symbol)&Gene_symbol %in% rownames(se)) %>%
    group_by(Gene_symbol)%>%
  slice(which.max(Overall_assoc.score))%>% #==> when there is a gene mentioned twice, only keep the row where the overall association score is the highest ==> reason for duplicates is that multiple human genes have the same mouse homolog etc...
    ungroup())

saveRDS(OT.files, file = paste("OpenTargets", sample,"OT_files_list.rds", sep = "/"))
```

# create gene lists

## newer approach: split OT.fles into genes that have scores for overall.assoc (optional), genetic.assoc and drug association
```{r}
# put this function in lapply 
split.OT.files <- function(x, name, prefix = T){
  # print(x)
  # print(name)
  # overall.associations <-  x %>%
  #   arrange(desc(Overall_assoc.score)) %>%
  #   filter(!is.na(Overall_assoc.score)) %>%
  #   .$Gene_symbol
  genetic.associations <-  x %>%
    arrange(desc(Genetic_assoc.)) %>%
    filter(!is.na(Genetic_assoc.)) %>%
    .$Gene_symbol
  drug.associations <-  x %>%
    arrange(desc(Drugs)) %>%
    filter(!is.na(Drugs)) %>%
    .$Gene_symbol
  # print(lis[[x]])
  if (prefix){
    # disease <- names(x[[prefix]])
    elements <- c(#paste(name,"overall.associations", sep = "_"),
                  paste(name,"genetic.associations", sep = "_"),
                  paste(name,"drug.associations", sep = "_"))
    # print(elements)
    list <- list(#overall.associations,
                  genetic.associations,
                  drug.associations )
    names(list) <- elements

  } else{
    list <- list(#"overall.associations" = overall.associations,
               "genetic.associations" = genetic.associations, 
               "drug.associations" = drug.associations )
    }
  return(list)
}


# list structure is: disease1-c(overall.associations,genetic.associations, drug.associations)
                      # disease2-c(overall.associations,genetic.associations, drug.associations)
#to get:                disease1-c(disease1_overall.associations,disease1_genetic.associations, disease1_drug.associations)
                      # disease2-c(disease2_overall.associations,disease2_genetic.associations, disease2_drug.associations) 
# state: prefix = T


```



```{r}
OT.files.split <- map2(OT.files, names(OT.files), split.OT.files, prefix = T) 
# can also use imap apparently, where "imap(x, ...), an indexed map, is short hand for map2(x, names(x), ...) if x has names,"
```

### filter genes to those also present in the SE object and get rid of empty sets

#### list.seurat.filter function (from "LCA_OT_pathway_enrichment_JBO..")
```{r }

list.seurat.filter <- function (list, seurat.object){ #for a list of lists or vectors, remove those lists/vectors that are smaller than "size" ==> for addmodulescore you need genes that are also present in the object
  list <- sapply(list, function(x) x[x %in% rownames(seurat.object)])
  list <- list[sapply(list, function(x) length(x)>0)]
  return(list)
}
```

```{r}
OT.files.split <- lapply(OT.files.split, list.seurat.filter, seurat.object = se)
```



```{r}
saveRDS(OT.files.split, file = paste("OpenTargets", sample,"OT_files_split", sep = "/"))

```

### create overlapping windows
```{r, overlapping windows function2}

splitWithOverlap <- function(vec, seg.length, overlap) {
  starts = seq(1, length(vec), by=seg.length-overlap)
  ends   = starts + seg.length - 1
  ends[ends > length(vec)] = length(vec)

  lapply(1:length(starts), function(i) vec[starts[i]:ends[i]])
}
win.size <- 50
move.by <- 10
overlap <- win.size-move.by

windows.list1 <- lapply(OT.files.split, function(x) lapply(x, splitWithOverlap, seg.length = win.size, overlap = overlap))

```

```{r}
names(windows.list1$NAFLD$NAFLD_genetic.associations)
```


### Addmodulescore from list of list(?)
modified function from "pathway sets enrichment"
```{r}

modulescores.from.list3 <- function(window.list, seurat.object){
  
  for (i in c(1:length(window.list))){
    # disease: e.g. "NAFLD"
    for (j in c(1:length(window.list[[i]]))){
      #set: genetic association, drug association ==> those are lists of windows
      # print(j)
      num.windows <- length(window.list[[i]][[j]])
      cat(names(window.list[[i]][j]), "\t", num.windows, " windows will be added", "\r\n")
      
      windows <- window.list[[i]][[j]] # list of vector of gene names
      print(class(windows))
      names <- paste0(names(window.list[[i]][j]),"_rank_")
      print(names)
      
      seurat.object <- AddModuleScore(seurat.object, features = windows, ctrl = 100, name = names)
      
      # cat(i,"/",num.windows, "\t", window.name, "\r")
      
      # cat(window.name, "\r")
   
    # print(set)

    }
  }
  # print(paste(num.all.sets, " added to seurat object"))
  return(seurat.object)
}

```


```{r}
names(windows.list1[["NAFLD"]][["NAFLD_genetic.associations"]])

```

### get name of pathways for plotting or subsetting of se object
modified function from "pathway sets enrichment"
```{r }
get.module.names <- function(window.list){
  vec.set.names <- vector()
  for (i in c(1:length(window.list))){
    # disease: e.g. "NAFLD"
    for (j in c(1:length(window.list[[i]]))){
      #set: genetic association, drug association ==> those are lists of windows
      names <- paste0(names(window.list[[i]][[j]]), c(1:length(window.list[[i]][[j]])))
      vec.set.names <- c(vec.set.names, names)
    }
  }
  # print(paste(num.all.sets, " added to seurat object"))
  return(vec.set.names)
}
```



```{r, warning=FALSE, eval = F}
se <- modulescores.from.list3(windows.list1, se)

features_to_plot <- get.module.names(windows.list1)
# features_to_plot[150:200]
```

```{r, eval = F}
meta.data <- se[[]]
# maybe select only the windows, in case you don't want to change the rest of the meta.data ==> or you write the entire meta data to the file and when loading it you can still select only the sliding windows
# used option 2 ==> 

write.table(meta.data,
          file = paste0(sample.path,"/", sample,"_sliding_window_meta_data.tsv"),
          sep = "\t",
          col.names = T,
          row.names=T)
```



# NULL set permutation AddModuleScore

## load clusters/zones
```{r, eval = nullset}
seurat_clusters <- read.csv(paste0(sample.path, "/seurat_clusters.csv"), sep ="\t", header = T)
res.columns <- colnames(seurat_clusters)[startsWith(colnames(seurat_clusters), "SCT_snn_res.")]
res.columns <- res.columns[grepl("0.4|0.8", res.columns)]
cur.res <- 0.8
cur.res <- paste0("SCT_snn_res.",cur.res)

se <- AddMetaData(se, seurat_clusters)
se$seurat_clusters <- se[[]][,cur.res]
```

## settings
```{r, eval = T}
# Run settings
permutation.nr <- 10000
genes.in.NULL.set <- 50

NULL.gene.set <- list()

for (i in 1:permutation.nr){
  genes <- sample(rownames(se), size = genes.in.NULL.set)
  NULL.gene.set[[i]] <-  genes
}
```

## addmodulescore
```{r, eval = T}
se.null.set <- se

se.null.set <- AddModuleScore(se.null.set, features = NULL.gene.set, name = "NULL_gene_set_")
```


```{r write raw null_set_data to file, eval = T}
#write raw null_set_data to file
se.null.set <- se.null.set[[]]
se.null.set <- se.null.set %>%
  select(contains("NULL_gene_set_"))
#write raw null_set_data to file
write.table(se.null.set, file = paste0(sample.path, "/se_null_set_meta_data", sample,"_g",genes.in.NULL.set,".tsv"),
          sep = "\t",
          row.names = T,
          col.names = T)
```





```{r, further process, eval = nullset}
#further process, calc median:
se.null.set <- se.null.set %>%
  summarise(across(contains("NULL_gene_set_"), median), .by = seurat_clusters)%>%
  column_to_rownames("seurat_clusters") %>%
  t() %>%
  as.data.frame()
# se_test2 <- reshape(se_test2, idvar = "seurat_clusters", direction = "wide") #necessary? cant remember

```


```{r, eval = nullset}
#write summarized null_set_data to file
write.table(se.null.set,
            file = paste0(sample.path, "/se_null_set_", sample,"_g", genes.in.NULL.set,".tsv"),
            sep = "\t",
            row.names = T,
            col.names = T
            )
# ==> file se_null_set_JBO009.tsv is written to folder livercellatlas_data/visium/mouse/GSM5764415
```


Session info
```{r}
sessionInfo()
```

