---
title: "LCA_sliding_window_JBO002_with_pvalue"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "reports")})
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    dev: 'CairoPNG'
date: "`r Sys.Date()`"
---

Load libraries:

```{r, message=FALSE}
library(STutility)
library(dplyr)
library(tidyverse)
library(data.table)
library(reshape2)
library(readxl)
library(ggplot2)
library(Seurat)
library(RColorBrewer)
```

# setup


```{r setup, include=FALSE}
library(knitr)

# which plots to print:
violin <- F
lineplots <- T
p.plots <- T
gridarrange <- F


# plot setups
# violin.w <- 12 
# violin.h <- 9
# 
# working with 4K monitor?
big_screen <- F

# global chunk settings
opts_chunk$set(echo = TRUE)
opts_chunk$set(dev = "CairoPNG")
opts_chunk$set(error = FALSE)# when FALSE, interrupt in case of errors                
opts_chunk$set(out.width = "100%")
opts_chunk$set(out.height = "100%")

#<<featureplot>>
opts_hooks$set(
# eval
  eval = function(options) {
    raw.code <- knit_code$get(opts_current$get()$label) 
    raw.code <- paste(raw.code, collapse = "")
    if (grepl("p.val.nullset|p_value|p.val|top.genetic.genes|p.value|<<calculate p-value>>|<<p-val-plot>>|<<load_p_val_nullset>>|<<addmodulescore_top_genes>>|<<spatial_p_plot>>", raw.code)) {
      options$eval = p.plots
      }
    if (grepl("<<line plot>>|line.plot.setup", raw.code)){
      options$eval = lineplots
      }
    if (grepl("VlnPlot|geom_violin", raw.code)){
      options$eval = violin
      }
    if (grepl("<<featureplot>>", raw.code)){
      options$eval = gridarrange
      }
  options
  })

opts_hooks$set(
# include
  include = function(options) {
    raw.code <- knit_code$get(opts_current$get()$label) 
    raw.code <- paste(raw.code, collapse = "")
    if (grepl("p.val.nullset|p_value|p.val|top.genetic.genes|p.value|<<calculate p-value>>|<<p-val-plot>>|<<load_p_val_nullset>>|<<addmodulescore_top_genes>>|<<spatial_p_plot>>", raw.code)) {
      options$include = p.plots
      }
    if (grepl("<<line plot>>|line.plot.setup", raw.code)){
      options$include = lineplots
      }
    if (grepl("VlnPlot|geom_violin", raw.code)){
      options$include = violin
      }
    if (grepl("<<featureplot>>", raw.code)){
      options$include  = gridarrange
    }
    
  options
  })

opts_hooks$set(
# fig.width
  fig.width = function(options) {
    raw.code <- knit_code$get(opts_current$get()$label) 
    raw.code <- paste(raw.code, collapse = "")
    if (grepl("p.plot.setup", raw.code)) {
        options$fig.width = 12
    }
    if (grepl("<<spatial_p_plot>>", raw.code)) {
        options$fig.width = 5
    }
    if (grepl("line plot|line.plot.setup", raw.code)){
        options$fig.width = 30
        }
    if (grepl("VlnPlot|geom_violin", raw.code)){
        options$fig.width = 12 
      }
    if (grepl("<<featureplot>>", raw.code)){
      options$fig.width  = 20
      }

    options
  })
opts_hooks$set(
# fig.height
  fig.height = function(options) {
    raw.code <- knit_code$get(opts_current$get()$label) 
    raw.code <- paste(raw.code, collapse = "")
    if (grepl("p.plot.setup", raw.code)) {
        options$fig.height = 9
    }
    if (grepl("<<spatial_p_plot>>", raw.code)) {
        options$fig.height = 4
    }
    if (grepl("line plot|line.plot.setup", raw.code)){
      options$fig.height = 30
      }
    if (grepl("VlnPlot|geom_violin", raw.code)){
      options$fig.height = 9
      }
    if (grepl("<<featureplot>>", raw.code)){
      options$fig.height  = 15
    }
    
    options
  }
)

opts_hooks$set(fig.height = function(options) {
  if (big_screen == T) {
    options$fig.height = options$fig.height*1.5
  }
  options
})


```


# Load ST data:

```{r}
infoTable <- read.csv("livercellatlas_data/visium/mouse/infoTable.csv", header = TRUE, sep =",")
sample <- "JBO002"
sample.path <- paste("livercellatlas_data/visium/mouse", infoTable[infoTable$short_file_name == sample,]$geo_accession, sep = "/")
se <-  InputFromTable(infotable = infoTable[infoTable$short_file_name == sample,],
                      minUMICountsPerGene = 100,
                      minSpotsPerGene = 5,
                      minUMICountsPerSpot = 500,
                      platform =  "Visium")
# num_samples <- length(unique(se@meta.data[["condition"]]))
img.save.path <- r"(C:\Users\aaronko\OneDrive - Karolinska Institutet\Uni\Master\4. Semester\thesis_project\writing\images\sliding_window\ST\JBO002\)"

```

Load image:

```{r}
se <- LoadImages(se, time.resolve = FALSE)
```

==\> note that this is only applicable to 1 samples. for more samples it has to be modified

```{r fig.width = 10, fig.height = 8}
p1 <- ggplot() +
  geom_histogram(data = se[[]], aes(nFeature_RNA), fill = "red", alpha = 0.7, bins = 50) +
  ggtitle("Unique genes per spot")

p2 <- ggplot() +
  geom_histogram(data = se[[]], aes(nCount_RNA), fill = "red", alpha = 0.7, bins = 50) +
  ggtitle("Total counts per spots")

gene_attr <- data.frame(nUMI = Matrix::rowSums(se@assays$RNA@counts), 
                        nSpots = Matrix::rowSums(se@assays$RNA@counts > 0))
p3 <- ggplot() +
  geom_histogram(data = gene_attr, aes(nUMI), fill = "red", alpha = 0.7, bins = 50) +
  scale_x_log10() +
  ggtitle("Total counts per gene (log10 scale)")

p4 <- ggplot() +
  geom_histogram(data = gene_attr, aes(nSpots), fill = "red", alpha = 0.7,  bins = 50) +
  ggtitle("Total spots per gene")

(p1 - p2)/(p3 - p4)
```

View image:

```{r}
ImagePlot(se, method = "raster")
```

Normalize data: se \<- SCTransform(se,return.only.var.genes = T)

```{r, message=FALSE, include=FALSE}
se <- SCTransform(se,return.only.var.genes = T)
```

FeatureOverlay

```{r, fig.width=8, fig.height=4}

spot_alpha = 0.2
point.size = 1.2

FeatureOverlay(se, 
               features = c("nCount_RNA","nFeature_RNA"),
               pt.alpha = spot_alpha,
               pt.size = point.size,
               add.alpha=T,
               ncols = 2,
               cols = c("black", "darkblue", "cyan", "yellow", "red", "darkred"))
```

## load zone and annotated seurat_clusterss from "spatial_structures" file

```{r, load clusters and zones}
res <- 0.8
seurat_clusters <- read.csv(paste0(sample.path, "/seurat_clusters_res",res, ".csv"), sep ="\t", header = T)
se$seurat_clusters <- seurat_clusters
se$seurat_clusters <-as.factor(se$seurat_clusters)

seurat_zones_df <- read.csv(paste(sample.path,"seurat_clusters_annot_zones.csv",sep = "/"), sep ="\t", header = T) 
seurat_zones <- seurat_zones_df$seurat_clusters
names(seurat_zones) <- row.names(seurat_zones_df)
seurat_zones <- sort(seurat_zones)
se$zone <- as.factor(seurat_zones)
```

```{r}
library(viridis)
n <- 18
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
col_vector[6] <- "white"
col_vector[8] <- "lightpink"
# col_vector2 <- c("darkred","orange","lightblue","darkblue",col_vector)
n <- length(unique(se$seurat_clusters))
col_vector2 <- turbo(n = n, begin = 0.2, end = 1)
se <- SetIdent(se, value="seurat_clusters")
```

## Cluster Feature Overlay

```{r,fig.width=10, fig.height=4}
# col_vector2 <- c("orange","lightblue","red","darkblue")

p1 <- FeatureOverlay(object = se, features = "zone", pt.alpha = 0.6, pt.size = 1.25, cols = col_vector2, pt.border = FALSE)
p2 <- FeatureOverlay(object = se, features = "seurat_clusters", pt.alpha = 0.6, pt.size = 1.25, cols = col_vector2, pt.border = FALSE)
p3 <- FeatureOverlay(object = se, features = "Cyp2e1", pt.alpha = 0.6, pt.size = 1.25, cols = c("black", "darkblue", "cyan", "yellow", "red", "darkred"), pt.border = FALSE)

p1|p2
p2|p3

```



# load other data

## load window meta data

from AddModuleScore (from before)

```{r}
meta.data <- read.csv(file =  paste(sample.path,
                       "JBO002_sliding_window_meta_data.tsv",
                       sep = "/"),
                      sep = "\t")
meta.data <- meta.data %>%
  dplyr::select(contains("_rank_"))
se@meta.data <- cbind(se@meta.data, meta.data) # combine both dataframes on the rownames
```

```{r}
meta.data %>%
  select(contains("_rank_")) %>%
  colnames() %>%
  gsub("_rank_.*","",.)%>%
  unique()

```

```{r, make summary df of the windows}
library("psych")
#here you can decide whether you wanna summarise by cluster or by zone! 
#zone or seurat_clusters
summarise.on <- "seurat_clusters"
meta.data <- se[[]]

summary.meta.clusters <- meta.data %>%
  dplyr::select(c(summarise.on, contains("_rank_")))

summary.meta.clusters <- describeBy(summary.meta.clusters, group = summary.meta.clusters[, summarise.on], quant = c(.05, .95)) 

summary.meta.clusters <- lapply(summary.meta.clusters, function(x) as.data.frame(x))


summary.meta.clusters <- map(summary.meta.clusters, ~.x %>%
                               rownames_to_column(., var = "window"))
summary.meta.clusters <- bind_rows(summary.meta.clusters, .id = summarise.on)
summary.meta.clusters <- summary.meta.clusters[-1,]
rownames(summary.meta.clusters) <- NULL
head(summary.meta.clusters)
```

```{r}
genetic.summary.meta.clusters <- summary.meta.clusters %>%
  filter(window %like% "genetic")
drug.summary.meta.clusters <- summary.meta.clusters %>%
  filter(window %like% "drug")

```

## load NULL set data

```{r, load NULL gene set data}

raw.NULL.set <- fread(list.files(sample.path, pattern = paste0("se_null_set_meta_data", sample,"_g50"), full.names=T ), sep = "\t") # for the line_plots ==> i.e. equivalent to "g50" 
gxx <- sub(".*_","",list.files(sample.path, pattern = paste0("se_null_set_meta_data", sample,"_g50")))
gxx <- sub(".tsv.*","",gxx)
colnames(raw.NULL.set)[1] <- "barcode"

raw.NULL.set$seurat_clusters <- NULL

raw.NULL.set <- raw.NULL.set %>%
  left_join(seurat_clusters %>% rownames_to_column(var = "barcode"), by = "barcode")

raw.NULL.set <- raw.NULL.set %>%
  summarise(across(contains("NULL_gene_set_"), median), .by = summarise.on)%>%
  column_to_rownames(summarise.on) %>%
  t() %>%
  as.data.frame()

NULL.set.data <- describe(raw.NULL.set, quant = c(.05, .95)) %>%
  as.data.frame %>%
  rownames_to_column(var = summarise.on) #summarized perm.matrix ==> median of all permutations


stacked.NULL.data <- stack(raw.NULL.set)
colnames(stacked.NULL.data) <- c("addmodulescore",summarise.on)
stacked.NULL.data$permutation <- rep(c(1:10000),length(unique(stacked.NULL.data[, summarise.on])))

remove(seurat_clusters)
```


# plotting
## introduce empty vector to store data
```{r for p value calculations, eval = T}
# not implemented yet, 11/04/23
important_plots <-  list()
```

## load OT.files

is easier to select the top genes and plot the density/p-value plot

```{r, code_folding = 'hide'}
OT.files.split <- readRDS(file = paste("OpenTargets", sample,"OT_files_split", sep = "/"))

OT.files.split.genetic <- sapply(OT.files.split, function(x) x[grep("genetic", names(x), value = T)])
names(OT.files.split.genetic) <- names(OT.files.split)
OT.files.split.genetic <- OT.files.split.genetic[sapply(OT.files.split.genetic, function(x) length(x)>0)]


OT.files.split.drug <- sapply(OT.files.split, function(x) x[grep("drug", names(x), value = T)])
names(OT.files.split.drug) <- names(OT.files.split)
OT.files.split.drug <- OT.files.split.drug[sapply(OT.files.split.drug, function(x) length(x)>0)]
OT.files.split.drug <- sapply(OT.files.split.drug, function(x) x[[1]])
```

## NAFLD {.tabset}

subset features_to_plot to specific disease for now

```{r}
disease <- "NAFLD"
features_to_plot <- se[[]] %>%
  dplyr::select(contains("_rank_")) %>%
  colnames()
features_to_plot <- features_to_plot[features_to_plot %like% disease]

```

### Violin Plots

```{r Violin Plots, eval = violin}
for (i in features_to_plot){

  plot(VlnPlot(se,
        features = i,
        ncol = 1,
        group.by = "seurat_clusters",
        cols = col_vector2,
        pt.size = 0.1) +
  geom_boxplot(width=0.4) +
  theme(plot.title = element_text(size = 12))
  )
}

```

### grid.arrange FeaturePlot

```{r featureplot, fig.width=20, fig.height=15, eval = gridarrange}
library(gridExtra)
max <- 12
num_plots <- c(1:length(features_to_plot))
num_plots <- split(num_plots, ceiling(seq_along(num_plots)/max))
spot.alpha <- 0.75
point.size = 2

for (i in num_plots){
  plot.list <- list()
  plots <- unlist(i)
  for (j in c(1:length(plots))){
    
    plot.list[[j]] <- 
      ST.FeaturePlot(se,
               features = features_to_plot[plots[j]],
               # pt.alpha = spot_alpha,
               pt.size = point.size,
               # grid.ncol = 4,
               cols = c("black", "darkblue", "cyan", "yellow", "red", "darkred"),
               value.scale = "all"
                )
  }
  do.call(grid.arrange, c(plot.list, ncol = 4, nrow = 3))
}

```

#### line plot setup

```{r, class.source = 'fold-hide'}

library(cowplot)
#boxplot on the right side
p.boxplot <- NULL.set.data %>%
  ggplot(aes(x = .data[[summarise.on]], ymin=min, lower= Q0.05, middle = median, upper = Q0.95, ymax = max)) +
  geom_boxplot(stat = "identity") +
  # ggtitle("")+
  theme_classic()+
  facet_wrap(vars(.data[[summarise.on]]), ncol = 1, scales = "free", strip.position = "left") 

line.plot.setup <- function(x){
  lwd <- 0.2
  lty <- 2
  p1 <- ggplot(x, aes(window, median, group = 1)) +
  geom_point(shape = "o", size = 2) +
  geom_line(linewidth = 0.1) +
  geom_hline(data = NULL.set.data, aes(yintercept  = min), colour = "darkred", lwd = lwd, lty = lty)+
  geom_hline(data = NULL.set.data, aes(yintercept  = max), colour = "darkred", lwd = lwd, lty = lty)+
  geom_hline(data = NULL.set.data, aes(yintercept  = median), col = "blue", lwd = lwd, lty = lty)+
  geom_hline(data = NULL.set.data, aes(yintercept  = Q0.05),color = "red", lwd = lwd, lty = lty)+
  geom_hline(data = NULL.set.data, aes(yintercept  = Q0.95), colour = "red", lwd = lwd, lty = lty)+
  xlab("Rank") +
  ylab("Score") +
  scale_x_discrete(labels = c(1:nrow(x)))+
  theme_classic()+
  facet_wrap(vars(.data[[summarise.on]]), ncol = 1, scales = "free")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank())
  
  
  combined.plot <- plot_grid(p.boxplot, p1,
                             align = "h", axis = "b", ncol = 2, rel_widths = c(2, 20)) #https://stackoverflow.com/questions/41834093/decrease-margins-between-plots-when-using-cowplot, not used but can be adjusted
  combined.plot
}
```

### line plots

```{r line plot, fig.width = 15, fig.height= 10}

genetic_features_to_plot <- summary.meta.clusters %>%
  filter(window %in% features_to_plot & window %like% "genetic") 
drug_features_to_plot <- summary.meta.clusters %>%
  filter(window %in% features_to_plot & window %like% "drug")

p <- line.plot.setup(genetic_features_to_plot) +
  ggtitle("genetic associations")

p

important_plots[[disease]][["line.plot"]] <- p
line.plot.setup(drug_features_to_plot) +
  ggtitle("drug associations")
# ggsave(filename = paste0(img.save.path, "line plot example NAFLD.svg"),
#        plot = p,
#        device = "svg",
#        dpi = 96,
#        height = 10,
#        width = 15)

```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 10
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r load_p_val_nullset}

p.val.nullset <- read.csv(paste0(sample.path, "/se_null_set_", sample,"_g", maximum,"_res",res,".tsv"), sep = "\t")
# 
# [] delete the "res" stuff when analysing the zones?
colnames(p.val.nullset) <- gsub("X","", colnames(p.val.nullset))
stacked.p.val.nullset <- stack(p.val.nullset)
colnames(stacked.p.val.nullset) <- c("addmodulescore","seurat_clusters")
stacked.p.val.nullset$permutation <- rep(c(1:10000),length(unique(stacked.p.val.nullset$seurat_clusters)))
```

##### addmodulescore for top genes

```{r addmodulescore_top_genes}
top.genetic.genes <- AddModuleScore(se, features = list(top.genetic.genes), name = paste0(disease, "_top_genetic_genes")) %>%
  .@meta.data %>%
  select(contains(paste0(disease, "_top_genetic_genes")), seurat_clusters)

top.genetic.genes <- top.genetic.genes %>%
  group_by(seurat_clusters)%>%
  summarise(across(where(is.numeric), median))%>% 
  column_to_rownames("seurat_clusters") %>%
  t() %>%
  as.data.frame()%>%
  stack()
colnames(top.genetic.genes) <- c("median","seurat_clusters")

```

##### calculate p-value

```{r calculate p-value}
p.value <- imap(p.val.nullset, function(x, idx){
  print(idx) #this is the column name
  median.to.compare.against <-  top.genetic.genes[top.genetic.genes$seurat_clusters == idx, "median"]
  print(paste0("median: ", median.to.compare.against))
  length(x[x>=median.to.compare.against]) #for each column/seurat_clusters in the raw.NULL.set, subset the column to those rows that have a higher median than the median of the top (70) genes for each cluster
})



p.value <- sapply(p.value, function(x) (x+1)/(10000+1)) 
# calculate the fraction of the medians of the  10000 permutations that are higher than the median of the top.genes median 
# think about p-value definition:
# “the probability of obtaining test results at least as extreme as the results actually observed, under the assumption that the null hypothesis is correct.” 
p.value.adj <- p.adjust(p.value, method = c("bonferroni")) #, n=length(unique(top.genetic.genes$seurat_clusters))
p.value.adj <- data.frame("p.value.adjusted" = p.value.adj, "p.val" = p.value ,"seurat_clusters" = as.factor(names(p.value.adj)))

top.genetic.genes <- top.genetic.genes %>%
left_join(p.value.adj) %>%
  arrange(p.value.adjusted)%>%
  mutate(significant = ifelse(p.value.adjusted<0.05,T,F),
         log10.p.adj = -log10(p.value.adjusted),
         log10.p = -log10(p.val),
         )
```

##### p-val plot setup

```{r p-val plot setup}
p.plot.setup <- function(top.genes.df){
  # adj_p_value <- ifelse(top.genes.df$p.value.adjusted<0.5,
  #                       top.genes.df$p.value.adjusted,
  #                       1-top.genes.df$p.value.adjusted)
  top.genes.df <- top.genes.df %>%
  arrange(p.value.adjusted)
  p <- ggplot(stacked.p.val.nullset, aes(x=addmodulescore)) +
  geom_density(fill = "#69b3a2",
               alpha = 0.8)+
  geom_vline(data = top.genes.df, aes(xintercept = median, color = significant),
             size = 1, show.legend = T) +
  scale_color_manual(values = c("yellow","red"), breaks = c("FALSE", "TRUE"))+
  geom_label(data = top.genes.df,
             aes(x=median+0.1, y = 8),
             label = paste0("adjusted p-value \n",
                      formatC(top.genes.df$p.value.adjusted,
                           format = "e", digits = 2)),
             size = 2) +
  xlab("Median score NULL gene set")+
  ylab("Density")
  
p +
  facet_wrap(~factor(seurat_clusters, levels = top.genes.df$seurat_clusters ))+
  theme_classic()
}
```

### p-value density plot

```{r p-val-plot, fig.height=9, fig.width=12}
p.plot.setup(top.genetic.genes)
ggsave(filename = paste0(img.save.path, "density p-val-plot NAFLD.svg"),
       device = "svg",
       dpi = 96,
       height = 9,
       width = 12)

important_plots[[disease]][["p.plot"]] <- p.plot.setup(top.genetic.genes)

```


purple line is the median addmodulescore of the top.genes in each cluster

### spatial p-val plot

```{r spatial_p_plot, fig.width=5, fig.height=4}
#<<spatial_p_plot>>
library(viridis)
p.vals <- se[[]]
p.vals<- p.vals %>%
  left_join(top.genetic.genes) %>%
  select(!contains("rank"))
rownames(p.vals) <- rownames(se[[]])
p.se <- se 
p.se <- AddMetaData(p.se, p.vals)
p <- ST.FeaturePlot(object = p.se,
              features = "log10.p",
              pt.size = 1,
              cols = col_vector2
              )+
  labs(fill="-log10(p-value)", title = "") 
p
# important_plots[[disease]][["spatial.p.plot"]] <- p
important_plots[[disease]][["top.genetic.genes.df"]] <- top.genetic.genes

ggsave(filename = paste0(img.save.path, "STFeaturePlot -log10 p-value NAFLD.pdf"),
       plot = p,
       device = "pdf",
       dpi = 96,
       height = 4,
       width = 5)
  
```

## liver_disease_general {.tabset}

subset features_to_plot to specific disease for now

```{r}
disease <- "liver_disease_general"

features_to_plot <- se[[]] %>%
  dplyr::select(contains("_rank_")) %>%
  colnames()
features_to_plot <- features_to_plot[features_to_plot %like% disease]
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```

### grid.arrange FeaturePlot

```{r, fig.width=16, fig.height=12, eval = gridarrange}
<<featureplot>>
```

### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 4
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p-value>>
```

### p-value density plot

```{r}
<<p-val-plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

### spatial p-val plot

```{r , fig.width=5, fig.height=4}
<<spatial_p_plot>>
```

## cirrhosis_of_liver {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "cirrhosis_of_liver"
features_to_plot <- se[[]] %>%
  dplyr::select(contains("_rank_")) %>%
  colnames()
features_to_plot <- features_to_plot[features_to_plot %like% disease]
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```

### grid.arrange FeaturePlot

```{r, fig.width=16, fig.height=12, eval = gridarrange}
<<featureplot>>
```

### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 10 #just all (107)
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p-value>>
```

### p-value density plot

```{r}
<<p-val-plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

### spatial p-val plot

```{r , fig.width=5, fig.height=4}
<<spatial_p_plot>>
```

## cirrhosis_familial {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "cirrhosis_familial"
features_to_plot <- se[[]] %>%
  dplyr::select(contains("_rank_")) %>%
  colnames()
features_to_plot <- features_to_plot[features_to_plot %like% disease]

```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```

### grid.arrange FeaturePlot

```{r, fig.width=16, fig.height=12, eval = gridarrange}
<<featureplot>>
```

### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 10
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p-value>>
```

### p-value density plot

```{r}
<<p-val-plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

### spatial p-val plot

```{r , fig.width=5, fig.height=4}
<<spatial_p_plot>>
```

## Diabetes type I {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "Diabetes_typeI"
features_to_plot <- se[[]] %>%
  dplyr::select(contains("_rank_")) %>%
  colnames()

features_to_plot <- features_to_plot[features_to_plot %like% disease]
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```

### grid.arrange FeaturePlot

```{r, fig.width=16, fig.height=12, eval = gridarrange}
<<featureplot>>
```

### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 15
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p-value>>
```

### p-value density plot

```{r}
<<p-val-plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

### spatial p-val plot

```{r , fig.width=5, fig.height=4}
<<spatial_p_plot>>
```

## Diabetes type II {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "diabetes_typeII"
features_to_plot <- se[[]] %>%
  dplyr::select(contains("_rank_")) %>%
  colnames()
features_to_plot <- features_to_plot[features_to_plot %like% disease]
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```

### grid.arrange FeaturePlot

```{r, fig.width=16, fig.height=12, eval = gridarrange}
<<featureplot>>
```

### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 21
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p-value>>
```

### p-value density plot

```{r}
<<p-val-plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

### spatial p-val plot

```{r , fig.width=5, fig.height=4}
<<spatial_p_plot>>
```

## triglyceride_measurement {.tabset}
subset features_to_plot to specific disease for now
OBS ==> there is also LDL_triglyceride_measurement, so have to exclude that 
```{r}
disease <- "triglyceride_measurement"

`%notlike%` <- negate(`%like%`)
features_to_plot <- se[[]] %>%
  dplyr::select(contains("_rank_")) %>%
  colnames()
features_to_plot <- features_to_plot[features_to_plot %like% disease]
features_to_plot <- features_to_plot[features_to_plot %notlike% "LDL"]
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```

### grid.arrange FeaturePlot

```{r, fig.width=16, fig.height=12, eval = gridarrange}
<<featureplot>>
```

### line plots

```{r, fig.width = 5, fig.height= 10}
<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 15
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p-value>>
```

### p-value density plot

```{r}
<<p-val-plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

### spatial p-val plot

```{r , fig.width=5, fig.height=4}
<<spatial_p_plot>>
```

## LDL_triglyceride_measurement {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "LDL_triglyceride_measurement"

features_to_plot <- se[[]] %>%
  dplyr::select(contains("_rank_")) %>%
  colnames()
features_to_plot <- features_to_plot[features_to_plot %like% disease]
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```

### grid.arrange FeaturePlot

```{r, fig.width=16, fig.height=12, eval = gridarrange}
<<featureplot>>
```

### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 6
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p-value>>
```

### p-value density plot

```{r}
<<p-val-plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

### spatial p-val plot

```{r , fig.width=5, fig.height=4}
<<spatial_p_plot>>
```

## cholesterol_measurement {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "cholesterol_measurement"

features_to_plot <- se[[]] %>%
  dplyr::select(contains("_rank_")) %>%
  colnames()
features_to_plot <- features_to_plot[features_to_plot %like% disease]
features_to_plot <- features_to_plot[features_to_plot %notlike% "LDL"]
features_to_plot <- features_to_plot[features_to_plot %notlike% "HDL"]

```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```

### grid.arrange FeaturePlot

```{r, fig.width=16, fig.height=12, eval = gridarrange}
<<featureplot>>
```

### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 16
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p-value>>
```

### p-value density plot

```{r}
<<p-val-plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

### spatial p-val plot

```{r , fig.width=5, fig.height=4}
<<spatial_p_plot>>
```

## LDL_cholesterol_measurement {.tabset}
subset features_to_plot to specific disease for now
```{r}
disease <- "LDL_cholesterol_measurement"

features_to_plot <- se[[]] %>%
  dplyr::select(contains("_rank_")) %>%
  colnames()
features_to_plot <- features_to_plot[features_to_plot %like% disease]
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```

### grid.arrange FeaturePlot

```{r, fig.width=16, fig.height=12, eval = gridarrange}
<<featureplot>>
```

### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 25
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p-value>>
```

### p-value density plot

```{r}
<<p-val-plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

### spatial p-val plot

```{r , fig.width=5, fig.height=4}
<<spatial_p_plot>>
```


## HDL_cholesterol_measurement {.tabset}
should be HDL_cholesterol
subset features_to_plot to specific disease for now
```{r, }
disease <- "HDL_cholesterol_measurement"

features_to_plot <- se[[]] %>%
  dplyr::select(contains("_rank_")) %>%
  colnames()
features_to_plot <- features_to_plot[features_to_plot %like% disease]
```

### Violin Plots

```{r , eval = violin}
<<Violin Plots>>
```

### grid.arrange FeaturePlot

```{r, fig.width=16, fig.height=12, eval = gridarrange}
<<featureplot>>
```

### line plots

```{r, fig.width = 5, fig.height= 10}

<<line plot>>
```

### p value permutation

#### top genetic genes

##### select top genes

```{r}
nr.top.genetic.windows <- 5
#what was the window size:
win.size <- 50
move.by <- 10
nr.top.genes <- win.size + ((nr.top.genetic.windows-1) * move.by)
maximum <- ifelse(length(OT.files.split.genetic[[disease]])<=nr.top.genes,
                  length(OT.files.split.genetic[[disease]]),
                  nr.top.genes) %>% as.numeric()
top.genetic.genes <- OT.files.split.genetic[[disease]][1:maximum]
```

#### load correct null set for p-value calculation

==> number of genes per permutation should equal the number of top.genetic.genes (nr.top.genes)

```{r}
<<load_p_val_nullset>>
```

##### addmodulescore for top genes

```{r}
<<addmodulescore_top_genes>>
```

##### calculate p-value

```{r}
<<calculate p-value>>
```

### p-value density plot

```{r}
<<p-val-plot>>
```

purple line is the median addmodulescore of the top.genes in each cluster

### spatial p-val plot

```{r , fig.width=5, fig.height=4}
<<spatial_p_plot>>
```


# save important plots
```{r}
object.size(important_plots)
saveRDS(important_plots,
        file = paste0(sample.path, "/important_plots_", sample,".rds"))
```

